---
title: "02 Null Model and ICC"
author: "Tami Bolodeoku & Carrick Davis"
date: "⁠ r Sys.Date() ⁠"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r load, include=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
library(here)

oasis <- read_csv(here("data", "raw", "oasis_longitudinal.csv"))
```

## Null Model

Before adding any predictors we fit a null model. This estimates the overall average brain volume and partitions variance between and within individuals. It has no predictors — it is simply asking: how much do people differ from each other in brain volume?


```{r null_model}
null_model <- lmer(nWBV ~ 1 + (1 | `Subject ID`), data = oasis)
summary(null_model)
```


## Intraclass Correlation Coefficient (ICC)

The ICC tells us what proportion of the total variance in brain volume is due to differences between people rather than changes within a person over time.

```{r}
var_components <- as.data.frame(VarCorr(null_model))
between_var <- var_components$vcov[1]
within_var  <- var_components$vcov[2]

icc <- between_var / (between_var + within_var)
cat("ICC =", round(icc, 3))
```